---
title: "Human Activity Project"
author: "Yiping Tang"
date: "Nov-11-2016"
output:
  html_document: default
  pdf_document: default
---

# Introduction
###  In this project, we will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. We will train models based on training dataset to predict in which manner they perform the barbell . Then we will test our models in test dataset to select our final model. Our data source has been provided by: http://groupware.les.inf.puc-rio.br/har. 


## Data Pre-processing
```{r}
Training<-read.csv("pml-training.csv",na.strings = c("","NA"))
Test<-read.csv("pml-testing.csv",na.strings = c("","NA"))
library(caret)
library(rattle)
library(randomForest)


Training<-Training[,c(-1,-3,-4)]
Test<-Test[,c(-1,-3,-4)]
set.seed(123)
split<-createDataPartition(Training$classe,p=0.8,list=FALSE)
train0<-Training[split,]
test0<-Training[-split,]

##Remove low variance variables 
nzv <- nearZeroVar(train0)
train0<-train0[,-nzv]
test0<-test0[,-nzv]

## Remove variables with lots of NA
list<-sapply(train0, function(x) mean(is.na(x))) > 0.5
train0<-train0[,list==F]
test0<-test0[,list==F]


```


## Model 1: Classification Tree
```{r}

## 10-fold Cross Validation
# We will split the training dataset into 10 fold, and train and test each fold to optimize and select the best model.
c<-trainControl(method = "cv", number =5)
model1<-train(classe ~ ., data = train0, method = "rpart", trControl = c)
fancyRpartPlot(model1$finalModel)

## Prediction
pre1<-predict(model1,test0)
confusionMatrix(test0$classe, pre1)

```
#### This model accuract is 0.54, and the expected out-of-sample error is 0.46. Generally, the performance of this classification tree is not good.


## Model 2: Random Forest

```{r}

model2<-randomForest(classe~.,data=train0)
  
pre2<-predict(model2,test0)
confusionMatrix(test0$classe,pre2)
```
#### The Random Forest model accuracy is 0.999. The expected out-of-sample error is less than 0.001. This model performance is much better than the previous Decision Tree. We would select this Random Forest Model as our final model for prediction.


## Prediction on Test
```{r}
Test<-Test[,-nzv]
Test<-Test[,list==F]
levels(Test$cvtd_timestamp)<-levels(train0$cvtd_timestamp)
prediction<-predict(model2,Test)
prediction

```


